\thesistitle

### 2019-08-23 sections/experimentalsetup.tex
:)
Sander

\section{The Corpus}\label{chap:corpus}
For our experiments we use a large corpus of open source projects \cite{githubCorpus2013}. This corpus contains a set of Java projects from GitHub, selected by number of forks. The projects in this corpus were then de-duplicated manually. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

@@

\section{Tool Validation}
We have validated the correctness of CloneRefactor through unit tests and empirical validation. First, we created a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, contents and refactorability category (see Section \ref{chap:contextsetup} and \ref{sec:refactorabilitysetup}), to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context and their proposed refactoring.

@@

Because we cloned the projects (and do not use the sources provided in the corpus), they still included generated source code. Because we do not want to analyze generated Java files, we decided to further filter the corpus to only include projects with a conventional structure. For Maven, projects are recommended to put their production code in the folder structure \texttt{src/main/java}. We omitted Maven projects that do not adhere to this convention.

\subsection{Gathering Dependencies}
As a next step, we collect all dependencies of the downloaded software projects. Maven uses the following command to automate this process: \texttt{mvn dependency:copy-dependencies -DoutputDirectory=lib}. This process can fail if external dependencies are no longer available. We removed such projects for which we could not obtain all dependencies.

\subsection{Building an AST of all Java files}
To verify the correctness of all Java files in the projects, we used JavaParser \cite{tomassetti2017javaparser} to create an AST of every Java file in the \texttt{src/main/java} folder of the software projects. A very small set of projects had syntactical errors, which we removed from the corpus.

\subsection{Inspecting outliers}
We ran CloneRefactor over every project in the corpus and inspected the output for each project. Some projects gave unusual high numbers of projects of a certain size, which we manually inspected. Often these projects contained generated source files. We removed these projects from the corpus.

\subsection{Resulting corpus}
This procedure results in 2,267 Java projects including all their dependencies. The projects vary in size and quality. The total size of all projects is 14.210.357 lines (11,315,484 when excluding whitespace) over a total of 99,586 Java files. This is an average of 6,268 lines over an average of 44 files per project, 141 lines on average per file. The largest project in the corpus is \textit{VisAD} with 502,052 lines over 1,527.

\section{Minimum clone size}
In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, when clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones cannot improve maintainability when refactored.

\section{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability (see Section~\ref{sec:metrics}). For our experiments, we aggregate the deltas obtained for these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item Normalizing each obtained metric delta over all deltas obtained a metric in our dataset results in equally weighted scores.
\end{itemize}
We derived these assumptions partly from Heitlager et al~\cite{heitlager2007practical} and our own expert intuition. The validity of these metrics influences the validity of the conclusions we can draw from our data. Because of that, in our experiments we focus mainly on significant differences in maintainability to draw a conclusion.

We normalize each obtained metric delta using the ``Standard score'', which is calculated as follows:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metrics. This method works well for normalization of our data, because as we divide by the standard deviation, outliers do not influence the resulting scores much.

We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}

>>>>

\section{The Corpus}\label{chap:corpus}
For our experiments we use a large corpus of open source projects \cite{githubCorpus2013}. %Sander: Miss wil je hier zeggen dat het gemaakt is door Allamanis et al ofziets
This corpus contains a set of Java projects from GitHub, selected by number of forks. The projects in this corpus were de-duplicated manually. % Sander: Wat werd er ge-de-dupliceerd? de files? de projects?
% Sander: Kan ik de corpus ergens vinden?
This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.
% Sander:huh wat je zegt nu eerst dat hij useful is, en daarmee meteen dat we hem hebben veranderd. waarom hebben we die controlled set nodig?

@@

\subsection{Tool Validation}
% Sander: deze eerst zin is heel plotseling, en relateerd niet meteen heel duidelijk aan de rest eromheen
% Sander: Eigenlijk past deze hele subsectie niet echt in het corpus gedeelte
We have validated the correctness of CloneRefactor through unit tests and empirical validation. First, we created a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, contents and refactorability category (see Section \ref{chap:contextsetup} and \ref{sec:refactorabilitysetup}), % Sander: weet je zeker dat contain goede woord is in die zin?
to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context and their proposed refactoring.

@@

Because we cloned the projects (and do not use the sources provided in the corpus) % Sander:Het is hier niet duidelijk wat je nog met corpus bedoelt. Jouw corpus? die van allamanis?
, they still included generated source code. Because we do not want to analyze generated Java files, we decided to further filter the corpus to only include projects with a conventional structure. For Maven, projects are recommended to put their production code in the folder structure \texttt{src/main/java}. We omitted Maven projects that do not adhere to this convention. % Sander: weten we nu zeker dat er geen generated files meer in staan?

\subsection{Gathering Dependencies}
As a next step, we collect all dependencies of the downloaded software projects. Maven uses the following command to automate this process: \texttt{mvn dependency:copy-dependencies -DoutputDirectory=lib}. This process can % Sander: can fail of will fail?
fail if external dependencies are no longer available. We removed such projects for which we could not obtain all dependencies.

\subsection{Building an AST of all Java files}
To verify the correctness of all Java files in the projects, we used JavaParser \cite{tomassetti2017javaparser} to create an AST of every Java file in the \texttt{src/main/java} folder of the software projects. A very small set of projects had syntactical errors, which we removed from the corpus. % Sander:examples of syntactical error? why did we remove those -> they crash the javaparser

\subsection{Inspecting outliers}
We ran CloneRefactor over every project in the corpus and inspected the output for each project. Some projects gave unusual high numbers of projects % Sander: clones?
of a certain size, which we manually inspected. Often these projects contained generated source files. We removed these projects from the corpus.

\subsection{Resulting corpus}
This procedure results in 2,267 Java projects including all their dependencies. The projects vary in size and quality. The total size of all projects is 14.210.357 lines (11,315,484 when excluding whitespace) over a total of 99,586 Java files. This is an average of 6,268 lines over an average of 44 files per project, 141 lines on average per file. The largest project in the corpus is \textit{VisAD} with 502,052 lines over 1,527. % Sander:files?

\section{Minimum clone size}
In this study, we want to find out what thresholds will improve maintainability if clones by those thresholds are refactored. For example, when clones are very small, refactoring them might not improve maintainability since the detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones cannot improve maintainability when refactored. % Sander: Hoe ben je hier achter gekomen?

\section{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability (see Section~\ref{sec:metrics}). For our experiments, we aggregate the deltas obtained for these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability. % Sander: Dit is niet echt assumptie toch
  \item Normalizing each obtained metric delta over all deltas obtained a metric in our dataset results in equally weighted scores. % Sander: zin klopt niet (en snap hem ook niet)
\end{itemize}
We derived these assumptions partly from Heitlager et al~\cite{heitlager2007practical} and our own expert intuition. The validity of these metrics influences the validity of the conclusions we can draw from our data. Because of that, in our experiments we focus mainly on significant differences in maintainability to draw a conclusion. % Sander: miss wil je deze laatste twee zin weghalen om jezelf niet in je voet te schieten, als je het maar kan beantwoorden tijdens je defense

We normalize each obtained metric delta using the ``Standard score'', which is calculated as follows:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metric. This method works well for normalization of our data, because as we divide by the standard deviation, outliers do not influence the resulting scores much.

We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}
% Sander: miss mooier met Maintainability = ....

### 2019-08-23 sections/results.tex
Added context
Sander

\section{Clone types}\label{sec:clonetypeexperiments}
In this section, we look into the differences between the different clone types that were introduced in Chapter \ref{chap:clonetypes}.

\subsection{Found nodes}
In this section, we display our results regarding the differences in found nodes between clone type 1-3~\cite{roy2007survey} and type 1R-3R. Figure~\ref{fig:typeres} shows the number of cloned nodes found over the entire corpus for the different clone types (the word ``Type'' is abbreviated as ``T''). For type 2R clones, we set the variability threshold (see Section~\ref{sec:variabilitythreshold}) to 10\%, meaning that no more than 10\% of expressions may differ between cloned fragments. For type 3 and 3R clones, we allow a gap of 20\% the clones surrounding it.

\begin{figure}[H]
  \centering
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{img/TypeResults}
    \caption{Number of cloned nodes.}
\label{fig:typeres}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{img/DurationChart}
    \caption{Clone type performance.}
  \label{fig:performance}
  \end{minipage}
\end{figure}

\subsection{Performance}
To map the viability of refactoring-oriented clone types for large scale clone detection, we measured the duration of finding clones by the different clone types. Figure~\ref{fig:performance} shows the duration of detecting all clones in the corpus using CloneRefactor for different clone types. This figure shows the average duration of running CloneRefactor 10 times for a certain clone type. We ran these performance tests on the compute nodes of the DAS4 supercomputer \cite{bal2016medium}, which do not have many background processes severely affecting the results. The durations are of course dependent on our implementation of the clone types.

@@

\subsection{Relation} \label{sec:relationresults}
Table~\ref{tab:relation} displays the number of clone classes found for the entire corpus for different relations (see Section~\ref{sec:setuprelation}).

@@

\subsection{Location}
Table~\ref{tab:contents} displays the number of clone classes found for the entire corpus for different locations (see Section~\ref{sec:setuplocation}).
\begin{table}[H]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Category & Clone instances & \% \\ \midrule
Method Level & 232,545 & 78.43\% \\
Class Level & 50,402 & 17.00\% \\
Constructor Level & 10,039 & 3.39\% \\
Interface Level & 2,693 & 0.91\% \\
Enum Level & 788 & 0.27\% \\
\end{tabular}
\caption{Amount of clone instances with a certain location}
\label{tab:location}
\end{table}

@@

\section{Extract Method}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table~\ref{tab:refactorability} shows to what extent clone classes can be refactored by using the ``Extract Method'' refactoring technique. The second column shows our measurements for the complete systems (just like the former experiments). The third column shows our measurements when restricting our search to method bodies. The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).

\begin{table}[H]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
\textit{\textbf{Category}} & \textit{\textbf{All}} & \textit{\textbf{\% (All)}} & \textit{\textbf{Method Body}} & \textit{\textbf{\% (Method Body)}} \\ \midrule
Can Be Extracted & 24,157 & 28.2\% & 26,109 & 50.4\% \\
Is Not A Partial Method & 21,625 & 25.3\% & 0 & 0.0\% \\
Top-level AST-Node is not a Statement & 19,887 & 23.2\% & 4,607 & 8.9\% \\
Spans Part of a Block & 12,964 & 15.2\% & 13,460 & 26.1\% \\
Multiple Return Values & 5,622 & 6.6\% & 6,131 & 11.9\% \\
Complex Control Flow & 1,106 & 1.3\% & 1,216 & 2.4\% \\
Overlap In Clone Class & 147 & 0.2\% & 147 & 0.3\% \\
Not In Class Or Interface & 70 & 0.1\% & 93 & 0.2\% \\
\bottomrule
\end{tabular}
\caption{Number of clones that can be extracted using the ``Extract Method'' refactoring technique}
\label{tab:refactorability}
\end{table}

\section{Refactoring}
%I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability, and gap size.
In our corpus, CloneRefactor has refactored 12,710 clone classes and measured the change in indicated metrics (see Section~\ref{sec:metrics}). Using the presented formulas (see Section~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Section~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring.

\subsection{Clone Token Volume}
Figure \ref{fig:maintainabilityscore} shows the obtained results when plotting the clone volume vs the maintainability increase/decrease. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. The maintainability score, as shown in Figure \ref{fig:maintainabilityscore}, is the average delta maintainability score of all clone classes with the specified token volume that are refactored. As the amount of data points decreases, the datapoints gain less statistical significance, because of which we represent the x-axis as a logaritmic scale. The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.

>>>>

\section{Clone types}\label{sec:clonetypeexperiments}
In this section, we look into the differences between the clone types that were introduced in Chapter \ref{chap:clonetypes}.

\subsection{Found nodes}
In this section, we display our results regarding the differences in found nodes between clone type % Sander:waar komt nodes ineens vandaan? ik ga uit van AST nodes maar is niet duidelijk denk ik
1-3~\cite{roy2007survey} and type 1R-3R. Figure~\ref{fig:typeres} shows the number of cloned nodes found over the entire corpus for the different clone types (the word ``Type'' is abbreviated as ``T''). For type 2R clones, we set the variability threshold (see Section~\ref{sec:variabilitythreshold}) to 10\%, meaning that no more than 10\% of expressions may differ between fragments to be considered clones. For type 3 and 3R clones, we allow a gap of 20\% the clones surrounding it.

\begin{figure}[H]
  \centering
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{img/TypeResults}
    \caption{Number of cloned nodes.}
\label{fig:typeres}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{img/DurationChart}
    \caption{Clone type performance.}
  \label{fig:performance}
  \end{minipage}
\end{figure}
% Sander: korte uitleg over wat we zien in de figures hier? (geld voor alle experimenten)

\subsection{Performance}
To map the viability of refactoring-oriented clone types for large scale clone detection, we measured the duration of finding clones by the different clone types. Figure~\ref{fig:performance} shows the duration of detecting all clones in the corpus using CloneRefactor for different clone types. This figure shows the average duration of running CloneRefactor 10 times for a certain clone type. We ran these performance tests on the compute nodes of the DAS4 supercomputer \cite{bal2016medium}, which do not have many background processes severely affecting the results. % Sander: zou zin na de komma (which..results) weglaten.
The durations are of course dependent on our implementation of the clone types.
% Sander: deze zin voegt ook niet zoveel toe

@@

\subsection{Relation} \label{sec:relationresults}
Table~\ref{tab:relation} displays the number of clone classes found for the entire corpus for different relations (see Section~\ref{sec:setuprelation}).
% Sander: is dat plaatje met overzicht van relaties van clone classen weg of ergens anders? hij is hier ook wel handig

@@

\subsection{Location}
Table~\ref{tab:contents} displays the number of clone classes found for the entire corpus for different locations (see Section~\ref{sec:setuplocation}).
\begin{table}[H]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Category & Clone instances & \% \\ \midrule
Method Level & 232,545 & 78.43\% \\
Class Level & 50,402 & 17.00\% \\
Constructor Level & 10,039 & 3.39\% \\
Interface Level & 2,693 & 0.91\% \\
Enum Level & 788 & 0.27\% \\
\end{tabular}
\caption{Amount of clone instances with a certain location} % Sander: miss: Amount of clone instances per location category ofzoiets
\label{tab:location}
\end{table}

@@

\section{Extract Method}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table~\ref{tab:refactorability} shows to what extent clone classes can be refactored by using the ``Extract Method'' refactoring technique. The second column shows our measurements for the complete systems (just like the former experiments). The third column shows our measurements when restricting our search to method bodies. % Sander: waarom willen we dat weten?
The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).
% Sander: die laatste zin snap ik niet
\begin{table}[H]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
\textit{\textbf{Category}} & \textit{\textbf{All}} & \textit{\textbf{\% (All)}} & \textit{\textbf{Method Body}} & \textit{\textbf{\% (Method Body)}} \\ \midrule
Can Be Extracted & 24,157 & 28.2\% & 26,109 & 50.4\% \\
Is Not A Partial Method & 21,625 & 25.3\% & 0 & 0.0\% \\
Top-level AST-Node is not a Statement & 19,887 & 23.2\% & 4,607 & 8.9\% \\
Spans Part of a Block & 12,964 & 15.2\% & 13,460 & 26.1\% \\
Multiple Return Values & 5,622 & 6.6\% & 6,131 & 11.9\% \\
Complex Control Flow & 1,106 & 1.3\% & 1,216 & 2.4\% \\
Overlap In Clone Class & 147 & 0.2\% & 147 & 0.3\% \\
Not In Class Or Interface & 70 & 0.1\% & 93 & 0.2\% \\
\bottomrule
\end{tabular}
\caption{Number of clones that can be extracted using the ``Extract Method'' refactoring technique}
\label{tab:refactorability}
\end{table}

\section{Refactoring}
%I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability, and gap size.
In our corpus, CloneRefactor has refactored 12,710 clone classes and measured the change in indicated metrics (see Section~\ref{sec:metrics}). Using the presented formulas (see Section~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Section~\ref{sec:characteristics}) % Sander: Ref geeft ???
influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring.

\subsection{Clone Token Volume}
Figure \ref{fig:maintainabilityscore} shows the obtained results when plotting the clone volume vs the maintainability increase/decrease. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. The maintainability score, as shown in Figure \ref{fig:maintainabilityscore}, is the average delta maintainability score of all clone classes with the specified token volume that are refactored. As the amount of data points decreases, the datapoints gain less statistical significance, because of which we represent the x-axis as a logaritmic scale.% Sander: weet niet zeker of dit logaritmische schaal heet
The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.
